---
title: "Presence of Stigma of Environmentally-Friendly Cloth Pads in India"
author: "Balqies Mohamed"
description: "Does Age and Stigma Towards Cloth, Reusable Menstrual Pads Have A Relationship?" 
output:
  distill::distill_article:
    self_contained: false
---



```{r}
library(tidyverse)
library(infer)

periods <- read_csv("Informed Choice for Better Menstrual Outcomes 2.csv")

periods<-periods|>
  mutate(person = if_else(NoStigma==1, 1, 1))
periods

```





FINAL PROJECT 

#Introduction

Amid increase environmental pollution and concern, researchers are looking for methods to reduce non-biodegradable waste. One of the sources of non-biodegradable waste is single use menstrual products. Alternatives to these single use products exist, yet many women still prefer single use menstrual products. By analyzing the cause of this preference, environmental groups can better encourage the uptake of reusable cloth pads. My research question asks: does age and stigma toward cloth, reusable menstrual pads have a relationship? 

I hypothesize that as age increases stigma towards the cloth menstrual products also increase. I hypothesize this because the reliable, reusable menstrual products are fairly new technology offered to women. Older women may be hesitant to re-use or try cloth pads to do their inefficiency in the past. If this is true, marketers of the product and environmental experts can better address this group's concerns. Especially in countries where waste management is weak, it is especially important to encourage uptake of reusable material. This is interesting because, as different demographics of women are encouraged to move towards environmentally-friendly menstrual products, plastic waste is able to be curbed by a significant amount. Tru Earth estimates that in one woman's lifetime using plastic pads, 28,000 pounds of waste is also created. 
https://www.tru.earth/Impact-of-Disposable-Feminine-Hygiene-Products-on-the-Environment#:~:text=In%20addition%2C%20the%20production%20process,than%2028%2C000%20pounds%20of%20waste.


# Data Section 

The data I use stems from research conducted by the University of Liverpool wherein they seek to examine the causes of certain attitudes regarding menstrual hygiene materials in India. They specifically examine attitudes toward using reusable cloth pads among women when provided with access to these cloth pads and information about them. While the data that I use for my project is not the outcomes of the cross-sectional study itself, but rather the recorded data prior to their application of the treatment (exposure to cloth pads), the University of Liverpool did use a series of cross-sectional suverys over the course six months. In this cross-sectional study, women of varying age groups within India where assigned to three different condition; the first group were instructed to use single use pads and given information. The second group was offered information on the benefits and drawbacks of different menstrual products and given cloth-pads. The third and final group only received the information about the pads. 


My independent variable is age. In figure 2, a graph visualizes the distribution of age among the participants in this study. As women tend to experience menopause around age 50, the number of women in that age group start to dwindle. Although age is not the dependent variable, it is helpful to view the spread of age to better understand the coefficient estimate later on. Age was recorded through self-reporting during the pre-survey before the actual experiment. 

My dependent variable is stigma. In a barplot labelled figure 1, I plot the spread of the presence of stigma among the participants.The participants recorded their stigma towards cloth pads as either present or not present, leading to stigma being marked as either 1 or 0. As seen in the graph, more than 150 participants feel as though they have no stigma towards cloth pads. Stigma exists in a little under 50 participants of the cross-sectional study. In the next section, I will present whether this stigma exists in a particular relationship with age. 


```{r}

IV <- periods |> 
  ggplot( mapping = aes( x = Age, y = Padusers)) + geom_col(fill = "darkseagreen4") + 
  labs( x = "Age of Particpant", y = "Number of Particpants", title = " Figure 2, Spread of Independent Variable: Participants by Age", caption = "Source: University of Liverpool")

IV 

```
My dependent variable is stigma. In a barplot labelled figure 1, I plot the spread of the presence of stigma among the participants.The participants recorded their stigma towards cloth pads as either present or not present, leading to stigma being marked as either 1 or 0. As seen in the graph, more than 150 participants feel as though they have no stigma towards cloth pads. Stigma exists in a little under 50 participants of the cross-sectional study. In the next section, I will present whether this stigma exists in a particular relationship with age. 

```{r}
periods<-periods|>
  mutate(nostig2= if_else(NoStigma==1, "No Stigma", "Stigma"))

DV <- periods |> 
  ggplot( mapping = aes( x = nostig2, y = Padusers)) + 
  geom_col(fill = "pink") + 
  labs( x  = "Presense of Stigma" , y =  "Number of Participants", 
        title = 
          "Figure 1, Spread of Dependent Variable: Presense of Stigma Amongst Participants")

DV
```
# Results
```{r}
##Barplot
periods<-periods|>
  mutate(nostig2= if_else(NoStigma==1, "No Stigma", 
                          "Stigma"))
periods |> 
  ggplot( mapping = aes( x = Age, y = person, 
                         fill = nostig2)) + 
  geom_col() + 
  labs( x = "Age of Particpant", 
        y = "Stigma Towards Cloth Reusable Pads", 
        title = "Presence of Stigma by Age", 
        caption = "Source: University of Liverpool") +
  scale_fill_discrete( name = "Stigma Status")

##Finding Proportion
 prop<- periods|>
  group_by(Age)|>
  summarize(prop = sum(NoStigma)/sum(person))
knitr::kable(prop)

## Regression
model<- lm(data = periods, NoStigma~Age)|>
  broom::tidy()|>
  select(term, est=estimate, se= std.error, pval=p.value)
knitr::kable(model, digits=3, col.names= c("Variable", 
                                           "Coefficient Estimate", 
                                           "Standard Error", "P Value"))


```

#Conclusion 

The results demonstrate that as age increases, the presence of stigma decreases. This proves my initial hypothesis incorrect as I had estimated that as age increase stigma would also increase. As seen by the graph and affirmed by the coefficient estimate, as age increases, the presence of stigma decreases. As seen by the graph and affirmed by the coefficient estimate, as age increases, the presence of stigma decreases. The proportions of stigma by age also offers an interesting perspective as some ages, for example 24 and 29, have no stigma whats so ever towards cloth pads. As the p-value is 0, we can conclude that these results are significant. 

One limitation of my analysis is that I do not include confounding variables in my analysis.For example, employment may be a confounding variable. Women who are unemployed, usually retired older women, may feel content with using less secure forms of menstrual products as they are in the comfort of their own home. In contrast, women in their twenties or thirties may feel extremely hesitant to use less secure forms of menstrual products, which ultimately adds to the stigma towards cloth pads as they are rumared to be less secure. Another limitation of my analysis is that there is a lack of information on socio-economic class. This may better explain the stigma of cloth pads. Cloth pads offer a financially conservative in comparison to splurging on plastic menstrual products. This may be utilized more by low-income women. As cloth pads may be viewed as used only by lower-class/ lower-cast individuals, the stigma of cloth pads may be present.  If I had more time/money, I would appreciate working in tangent with the University of Liverpool to collect more information of class/caste difference in use of menstrual products.As I am not India, I also like to ask India researchers  to gain more knowledge about the use of menstrual products and attitudes In India. It may be so that it is a norm that women over 30 transition to using cloth pads, while younger women use single use pads. If the results turn out to be a cultural phenomenon, I would like to study the correlation of age and stigma of cloth pads globally. As all women in every corner of the global menstruate, it would be valuable to discover which groups to target when encourgaing uptake of cloth pads for environmental reasons.


