---
title: "Gov 50 Final Project"
author: "Balqies Mohamed"
description: "A Project to Exam Period Product Waste and Waste Management in Developing and Low-income Countries" 
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to period waste. I am passionate about period poverty and education, specifically in low-income and developing countries. I want to answer how much period products, which are largely one-time use and non-biodegradable, women use in developing countries in relation to the efficiency and productivity of waste management in that country. Not only is this a question of sustainability, but also one of public health. 


## Proposal -- milestone 2 


Does access to reusable menstrual products change the attitudes of women towards reusable products? (does access cause uptake) To answer this question, I plan to look at a data set from an experiment in which women are interviewed before and after they were given re-usable cloth pads. I hypothesize that with legitimate access to re-usable pads, women will change their preferences in which period products they chose to use outside of the study.The theory behind this is that access makes people realize that one kind of period products, the reusable kind, are preferable. The reason people do not switch is lack of access. If this is true, we should see an increase in the community where it is accessible. The explanatory variables is the attitudes of women towards reusable pads before exposure to it and after. My outcome variable is the attitudes and uptake of reusable pads. This data will be measured through the use of 0 and 1, wherein 1 indicates a women has preferences towards reusable pads. If my hypothesis is true we would observe a positive significant coefficient between uptake and access. If my hypothesis is incorrect, we would obserce a negative correlation between uptake and access.

```{r}
library(tidyverse)
periods <- read_csv("Informed Choice for Better Menstrual Outcomes 2.csv")

periods<-periods|>
  mutate(person = if_else(NoStigma==1, 1, 1))
periods

```
##Milestone 3 

```{r}
periods |> 
  ggplot( mapping = aes( x = Age, y = Padusers)) + geom_col(fill = "darkseagreen4") + 
  labs( x = "Age of Particpant", y = "Number of Particpants Using Pads", title = "Number of Padusers by Age", caption = "Source: University of Liverpool")
```
## Milestone 4 

```{r}

library(broom)
library(tidyverse)

periods<-periods|>
  mutate(nostig2= if_else(NoStigma==1, "No Stigma", "Stigma"))
periods |> 
  ggplot( mapping = aes( x = Age, y = person, fill = nostig2)) + geom_col() + 
  labs( x = "Age of Particpant", y = "Stigma Towards Cloth Reusable Pads", title = "Presence of Stigma by Age", caption = "Source: University of Liverpool") + scale_fill_discrete( name = "Stigma Status")


 prop<- periods|>
  group_by(Age)|>
  summarize(prop = sum(NoStigma)/sum(person))
knitr::kable(prop)


model<- lm(data = periods, NoStigma~Age)|>
  broom::tidy()|>
  select(term, est=estimate, se= std.error, pval=p.value)

knitr::kable(model, digits=3, col.names= c("Variable", "Coefficient Estimate", "Standard Error", "P Value"))


```

In realizing my data could not answer my initial research question- does access to cloth pads cause uptake? I found a different, but equally interesting and helpful question to answer: does age and stigma towards cloth,reusable menstrual pads have a relationship? 

My first visualization titled "Number of PadUsers by Age" demonstrates how many pad users within an age group. While this does not directly answer my new research question, the data presented is  useful to determining who is most likely to use disposable pads. As seen, women over 40 are increasingly less likely to use disposible pads. This is not because of menopause as the study only includes women who have not yet reached menopause.Thus it may be concluded that women around the ages 39-60 are less likely to have been using disposable pads. 

In the second visualization titled "Presence of Stigma by Age" directly answers the question of relationship between a the presence of stigma towards cloth pads and age. As seen by the graph and affirmed by the coefficient estimate, as age increases, the presence of stigma decreases. The proportions of stigma by age also offers an interesting perspective as some ages, for example 24 and 29, have no stigma whats so ever towards cloth pads. 


